Unicode Implementation Issues \thanks { This is a working document for the Rexx Language Association (RexxLA). There will be one or more separate requirements documents. It is written in \LaTeX\ \fmtversion\ and was rendered \today\ using \enginedetails. }

### 2024-02-16 ARB/standards/work-in-progress/document/tex/standard/evaluation.tex
work in standard book production process including readme.md
upd

\hypertarget{evaluation}{%
\chapter{Evaluation}\label{evaluation}}

@@

\hypertarget{variables}{%
\section{Variables}\label{variables}}

@@

\hypertarget{var_empty}{%
\subsection{Var\_Empty}\label{var_empty}}

@@

\hypertarget{var-set}{%
\subsection{Var Set}\label{var-set}}

@@

\hypertarget{var_value}{%
\subsection{Var\_Value}\label{var_value}}

@@

\hypertarget{var_drop}{%
\subsection{Var\_Drop}\label{var_drop}}

@@

\hypertarget{var_expose}{%
\subsection{Var\_Expose}\label{var_expose}}

@@

\hypertarget{var_reset}{%
\subsection{Var\_Reset}\label{var_reset}}

@@

\hypertarget{symbols}{%
\section{Symbols}\label{symbols}}

@@

\hypertarget{value-of-a-variable}{%
\section{Value of a variable}\label{value-of-a-variable}}

@@

\hypertarget{value-of-a-reserved-symbol}{%
\subsection{Value of a reserved
symbol}\label{value-of-a-reserved-symbol}}

@@

\hypertarget{expressions-and-operators}{%
\section{Expressions and operators}\label{expressions-and-operators}}

@@

\hypertarget{the-value-of-a-term}{%
\subsection{The value of a term}\label{the-value-of-a-term}}

See nnn for the syntax of a ferm.

@@

\hypertarget{the-value-of-a-prefix_expression}{%
\subsection{The value of a
prefix\_expression}\label{the-value-of-a-prefix_expression}}

@@

\hypertarget{the-value-of-a-power_expression}{%
\subsection{The value of a
power\_expression}\label{the-value-of-a-power_expression}}

@@

\begin{verbatim}
if datatype(lhs)\=='NUM' then call #Raise 'SYNTAX',41.1,lhs,'**'

if \datatype(rhs,'W') then call #Raise 'SYNTAX',26.1,rhs,'**'
\end{verbatim}

The value of the power\_expression is

@@

\hypertarget{the-value-of-a-multiplication}{%
\subsection{The value of a
multiplication}\label{the-value-of-a-multiplication}}

@@

\hypertarget{the-value-of-an-addition}{%
\subsection{The value of an addition}\label{the-value-of-an-addition}}

@@

\hypertarget{the-value-of-a-concatenation}{%
\subsection{The value of a
concatenation}\label{the-value-of-a-concatenation}}

@@

\hypertarget{the-value-of-a-comparison}{%
\subsection{The value of a comparison}\label{the-value-of-a-comparison}}

@@

If the comparison has a comparison\_operator that is a normal\_compare
then the variable \#Test is set as follows:

\begin{verbatim}
if datatype(lhs)\== 'NUM' | datatype(rhs)\== 'NUM' then do
/* Non-numeric non-strict comparison */
lhs=strip(lhs, 'B', ' ') /* ExtraBlanks not stripped */
rhs=strip(rhs, 'B', ' ')

if length(lhs)>length(rhs) then rhs=left (rhs, length (lhs) )
else lhs=left (lhs, length (rhs) )
if lhs>>rhs then #Test='G'
else if lhs<<rhs then #Test='L'
else #Test='E'

end
else do /* Numeric comparison */
if left(-lhs,1) == '-' & left(+rhs,1) \== '-' then #Test='G!
else if left(-rhs,1) == '-' & left(+lhs,1) \== '-' then #Test='L'
else do
Difference=lhs - rhs /* Will never raise an arithmetic condition. */
if Difference > 0 then #Test='G'
else if Difference < 0 then #Test='L'
else #Test='E'
end
end
The value of #Test, in conjunction with the operator in the comparison, determines the value of the
comparison.
The value of the comparison is '1' if
- #Test is 'E' and the operator is one of '="", '=="", '>=', <=", ‘\>', '\<', 'p>=', '<<=', \>>', or <<)

- #Test is 'G' and the operator is one of '>', '>=", ‘\<', ‘\=', '<>', '><', Nes", ‘>>! ‘p>’, or <<")
- #Test is 'L' and the operator is one of '<', <=", \>', \=', '<>', '><', \==', '<<', *<<=', or \>>'.
In all other cases the value of the comparison is '0'.

If the comparison is not a concatenation then:
if #Tracing.#Level == 'I' then call #Trace '>0O>'
\end{verbatim}

\hypertarget{the-value-of-an-and_expression}{%
\subsection{The value of an
and\_expression}\label{the-value-of-an-and_expression}}

@@

\hypertarget{the-value-of-an-expression}{%
\subsection{The value of an
expression}\label{the-value-of-an-expression}}

@@

\hypertarget{arithmetic-operations}{%
\subsection{Arithmetic operations}\label{arithmetic-operations}}

@@

The code of ArithOp itself is assumed to operate under a sufficiently
high setting of numeric digits to avoid exponential notation.

\begin{verbatim}
ArithoOp:

arg Numberl, Operator, Number2
/* The Operator will be applied to Numberl and Number2 under the numeric
settings #Digits.#Level, #Form.#Level, #Fuzz.#Level */

/* The result is the result of the operation, or the raising of a 'SYNTAX' or
'LOSTDIGITS' condition. */

/* Variables with digit 1 in their names refer to the first argument of the
operation. Variables with digit 2 refer to the second argument. Variables
with digit 3 refer to the result. */

/* The quotations and page numbers are from the first reference in
Annex C of this standard. */

/* The operands are prepared first. (Page 130) Function Prepare does this,
separating sign, mantissa and exponent. */

v = Prepare (Numberl, #Digits.#Level)
parse var v Signl Mantissal Exponentl

v = Prepare (Number2, #Digits.#Level)
parse var v Sign2 Mantissa2 Exponent2

/* The calculation depends on the operator. The routines set Sign3
Mantissa3 and Exponent3. */

Comparator = ''

select
when Operator == '*' then call Multiply

when Operator
when Operator

' then call DivType
*' then call Power
when Operator %' then call DivType
when Operator '//" then call DivType
otherwise call AddSubComp

end

call PostOp /* Assembles Number3 */

if Comparator \== '' then do
/* Comparison requires the result of subtraction made into a logical */
/* value. */
t = '0'
select
when left (Number3,1) == '-' then
if wordpos(Comparator,'< <= <> >< \= \>') > 0 then t = '1'
when Number3 \== '0' then
if wordpos(Comparator,'> >= <> >< \= \<') > 0 then t = '1'
otherwise
if wordpos(Comparator,'>= = =< \< \>') > 0 then t = '1!
end
Number3 = t
end
return Number3 /* From ArithOp */
/* Activity before every operation: */
Prepare: /* Returns Sign Mantissa and Exponent */
/* Preparation of operands, Page 130 */
/* ",...terms being operated upon have leading zeros removed (noting the

position of any decimal point, and leaving just one zero if all the digits in

the number are zeros) and are then truncated to DIGITS+1 significant digits
(if necessary)..." */

arg Number, Digits

/* Blanks are not significant. */

/* The exponent is separated */

parse upper value space(Number,0) with Mantissa 'E' Exponent
if Exponent == '' then Exponent = '0'

/* The sign is separated and made explicit. */

Sign = '+' /* By default */
if left(Mantissa,1) == '-' then Sign = '-'
if verify (left (Mantissa,1),'+-') = 0 then Mantissa = substr(Mantissa,2)

/* Make the decimal point implicit; remove any actual Point from the
Mantissa. */
Pp = pos('.',Mantissa)
if p > 0 then Mantissa = delstr(Mantissa,p,1)
else p = 1+length (Mantissa)

/* Drop the leading zeros */
do q = 1 to length(Mantissa) - 1

if substr(Mantissa,g,1) \== '0' then leave
p=p-l
end q

Mantissa = substr(Mantissa,q)

/* Detect if Mantissa suggests more significant digits than DIGITS
caters for. */
do j = Digits+1 to length (Mantissa)
if substr(Mantissa,j,1) \== '0' then call #Raise 'LOSTDIGITS', Number
end j

/* Combine exponent with decimal point position, Page 127 */
/* "Exponential notation means that the number includes a power of ten

following an 'E' that indicates how the decimal point will be shifted. Thus

4E9 is just a shorthand way of writing 4000000000 " */

/* Adjust the exponent so that decimal point would be at right of
the Mantissa. */
Exponent = Exponent - (length(Mantissa) - p + 1)

/* Truncate if necessary */

t = length(Mantissa) - (Digits+1)

if t > 0 then do
Exponent = Exponent + t
Mantissa = left (Mantissa,Digits+1)
end

if Mantissa == '0' then Exponent = 0

return Sign Mantissa Exponent

/* Activity after every operation. */
/* The parts of the value are composed into a single string, Number3. */
PostOp:

/* Page 130 */
/* 'traditional' rounding */
t = length(Mantissa3) - #Digits.#Level
if t > 0 then do
/* 'traditional' rounding */
Mantissa3 = left (Mantissa3,#Digits.#Level+1) + 5
if length(Mantissa3) > #Digits.#Level+1 then
/* There was 'carry' */

Exponent3 = Exponent3 + 1

Mantissa3 = left (Mantissa3,#Digits.#Level)

Exponent3 = Exponent3 + t

end
/* "A result of zero is always expressed as a single character '0' "*/
if verify (Mantissa3,'0') = 0 then Number3 = '0'
else do

if Operator == '/' | Operator == '**' then do

/* Page 130 "For division, insignificant trailing zeros are removed
after rounding." */

/* Page 133 "... insignificant trailing zeros are removed." */
do q = length(Mantissa3) by -1 to 2
if substr(Mantissa3,q,1) \== '0' then leave
Exponent3 = Exponent3 + 1
end q
Mantissa3 = substr(Mantissa3,1,q)
end
if Floating() == 'E' then do /* Exponential format */

Exponent3 = Exponent3 + (length(Mantissa3)-1)

/* Page 136 "Engineering notation causes powers of ten to be expressed as a

multiple of 3 - the integer part may therefore range from 1 through
999." */

g=l

if #Form.#Level == 'E' then do

/* Adjustment to make exponent a multiple of 3 */
g = Exponent3//3 /* Recursively using ArithOp as
an external routine. */
if g < 0 then g =g+ 3
Exponent3 = Exponent3 - g
geaqgitl
if length(Mantissa3) < g then
Mantissa3 = left (Mantissa3,g,'0')
end /* Engineering */

/* Exact check on the exponent. */
if Exponent3 > #Limit ExponentDigits then

call #Raise 'SYNTAX', 42.1, Numberl, Operator, Number2
if -#Limit ExponentDigits > Exponent3 then

call #Raise 'SYNTAX', 42.2, Numberl, Operator, Number2

/* Insert any decimal [point. */

if length(Mantissa3) \= g then Mantissa3 = insert('.',Mantissa3,g)
/* Insert the E */
if Exponent3 >= 0 then Number3
else Number3
end /* Exponent format */
else do /* 'pure number' notation */
p = length(Mantissa3) + Exponent3 /* Position of the point within
Mantissa */
/* Add extra zeros needed on the left of the point. */
if p < 1 then do
Mantissa3 = copies('0',1 - p)| |Mantissa3
p=il
end
/* Add needed zeros on the right. */
if p > length(Mantissa3) then
Mantissa3 = Mantissa3||copies('0',p-length (Mantissa3) )
/* Format with decimal point. */
Number3 = Mantissa3

Mantissa3'E+'Exponent3
Mantissa3'E'Exponent3

if p < length(Number3) then Number3 = insert('.',Mantissa3,p)
else Number3 = Mantissa3
end /* pure */
if Sign3 == '-' then Number3 = '-'Number3
end /* Non-Zero */
return
/* This tests whether exponential notation is needed. */
Floating:
/* The rule in the reference has been improved upon. */
Ct = ter
if Exponent3+length(Mantissa3) > #Digits.#Level then t = 'E'
if length(Mantissa3) + Exponent3 < -5 then t = 'E'
return t
/* Add, Subtract and Compare. */

AddSubComp: /* Page 130 */
/* This routine is used for comparisons since comparison is
defined in terms of subtraction. Page 134 */
/* "Numeric comparison is affected by subtracting the two numbers (calculating

the difference) and then comparing the result with '0'." */
NowDigits = #Digits.#Level
if Operator \=='+' & Operator \== '-' then do

Comparator = Operator

/* Page 135 "The effect of NUMERIC FUZZ is to temporarily reduce the value
of NUMERIC DIGITS by the NUMERIC FUZZ value for each numeric comparison" */
NowDigits = NowDigits - #Fuzz.#Level

end
/* Page 130 "If either number is zero then the other number ... is used as
the result (with sign adjustment as appropriate). */
if Mantissa2 == '0' then do /* Result is the lst operand */
Sign3=Signl; Mantissa3 = Mantissal; Exponent3 = Exponentl
return ''
end

if Mantissal
Sign3 = Sign

== '0' then do /* Result is the 2nd operand */
2
if Operator \

; Mantissa3 = Mantissa2; Exponent3 = Exponent2
== '+' then if Sign3 = '+' then Sign3 rat
else Sign3

bat
return ''
end

/* The numbers may need to be shifted into alignment. */

/* Change to make the exponent to reflect a decimal point on the left,

so that right truncation/extension of mantissa doesn't alter exponent. */
Exponentl = Exponentl + length (Mantissal1)
Exponent2 = Exponent2 + length (Mantissa2)

/* Deduce the implied zeros on the left to provide alignment. */

Aligni = 0

Align2 = Exponentl - Exponent2

if Align2 > 0 then do /* Arg 1 provides a more significant digit */
Align2 = min(Align2,NowDigits+1) /* No point in shifting further. */
/* Shift to give Arg2 the same exponent as Argl */

Mantissa2 = copies('0',Align2) || Mantissa2
Exponent2 = Exponentl
end

if Align2 < 0 then do /* Arg 2 provides a more significant digit */
/* Shift to give Argl the same exponent as Arg2 */

Aligni = -Align2

Alignl = min(Align1l,NowDigits+1) /* No point in shifting further. */
Align2 = 0

Mantissal = copies('0',Alignl) || Mantissal

Exponentl = Exponent2

end

/* Maximum working digits is NowDigits+1. Footnote 41. */

SigDigits
SigDigits

max (length (Mantissal) , length (Mantissaz2) )
min (SigDigits,NowDigits+1)

/* Extend a mantissa with right zeros, if necessary. */
Mantissal = left (Mantissal,SigDigits,'0')

Mantissa2 = left (Mantissa2,SigDigits,'0')

/* The exponents are adjusted so that

the working numbers are integers, ie decimal point on the right. */

Exponent3 = Exponentl-SigDigits
Exponentl = Exponent3
Exponent2 = Exponent3
if Operator = '+' then
Mantissa3 = (Signl || Mantissal) + (Sign2 || Mantissa2)

else Mantissa3 (Signl || Mantigsal) - (Sign2 || Mantissa2)

/* Separate the sign */
if Mantissa3 < 0 then do

Sign3 = '-'
Mantissa3 = substr (Mantissa3,2)
end

else Sign3 = '+'

/* "The result is then rounded to NUMERIC DIGITS digits if necessary,
taking into account any extra (carry) digit on the left after addition,
but otherwise counting from the position corresponding to the most
significant digit of the terms being added or subtracted." */

if length(Mantissa3) > SigDigits then SigDigits = SigDigits+l
d = SigDigits - NowDigits /* Digits to drop. */
if d <= 0 then return
t = length(Mantissa3) - d /* Digits to keep. */
/* Page 130. "values of 5 through 9 are rounded up, values of 0 through 4 are
rounded down." */
if t > 0 then do
/* 'traditional' rounding */
Mantissa3 = left(Mantissa3, t +1) +5
if length(Mantissa3) > t+1 then
/* There was 'carry' */
/* Keep the extra digit unless it takes us over the limit. */
if t < NowDigits then t = t+l
else Exponent3 = Exponent3+1
Mantissa3 = left (Mantissa3,t)
Exponent3 = Exponent3 + d
end /* Rounding */
else Mantissa3 = '0'
return /* From AddSubComp */

/* Multiply operation: */

Multiply: /* p 131 */

/* Note the sign of the result */

if Signl == Sign2 then Sign3 = '+'
else Sign3 = '-'
/* Note the exponent */
Exponent3 = Exponentl + Exponent2
if Mantissal == '0' then do
Mantissa3 = '0'
return
end
/* Multiply the Mantissas */
Mantissa3 = ''

do q=1 to length (Mantissa2)
Mantissa3 = Mantissa3'0'
do substr(Mantissa2,q,1)
Mantissa3 = Mantissa3 + Mantissal
end
end q
return /* From Multiply */

/* Types of Division: */

DivType: /* p 131 */
/* Check for divide-by-zero */

if Mantissa2 == '0' then call #Raise 'SYNTAX',

/* Note the exponent of the result */
Exponent3 = Exponentl - Exponent2

/* Compute (one less than) how many digits will be in the integer

part of the result. */

IntDigits = length(Mantissal) - Length(Mantissa2) + Exponent3
/* In some cases, the result is known to be zero.
if Mantigsal = 0 | (IntDigits < 0 & Operator

Mantissa3 = 0
Sign3 = '+'
Exponent3 = 0
return
end
/* In some cases, the result is known to be to be the first argument.
if IntDigits < 0 & Operator == '//' then do
Mantissa3 = Mantissal
Sign3 = Signl
Exponent3 = Exponentl
return
end
/* Note the sign of the result. */
if Signl == Sign2 then Sign3 = '+'
else Sign3 = '-'

/* Make Mantissal at least as large as Mantissa2 so Mantissa2 can be
subtracted without causing leading zero to result. Page 131 */

az 0
do while Mantissa2 > Mantissal

Mantissal = Mantissal'0'
Exponent3 = Exponent3 - 1
aztadtl
end
/* Traditional divide */
Mantissa3 = ''

/* Subtract from part of Mantissal that has length of Mantissa2 */

left (Mantissal,length(Mantissa2) )
substr (Mantissal, length (Mantissa2)+1)
o forever

x
Y
d

/* Develop a single digit in z by repeated subtraction.

ze=O0
do forever
xX = kK - Mantissa2
if left(x,1) == '-' then leave
Zeze¢t+tl
end

x = x + Mantissa2 /* Recover from over-subtraction */
/* The digit becomes part of the result */

Mantissa3 = Mantissa3 || z

if Mantissa3 == '0' then Mantissa3 = '' /* A single leading

##

*f
'%') then do

*/
zero can happen. */
/* x||y is the current residue */
if y == '' then if x = 0 then leave /* Remainder is zero */

if length(Mantissa3) > #Digits.#Level then leave /* Enough digits
in the result */

/* Check type of division */
if Operator \== '/' then do
if IntDigits = 0 then leave
IntDigits = IntDigits - 1
end
/* Prepare for next digit */
/* Digits come from y, until that is exhausted. */
/* When y is exhausted an extra zero is added to Mantissal */
if y == '' then do
y = ror
Exponent3 = Exponent3 - 1
aztadtl
end
xX = xX | | left (y,1)
y = substr(y,2)
end /* Iterate for next digit. */

Remainder = x || y
Exponent3 = Exponent3 + length(y) /* The loop may have been left early.
/* Leading zeros are taken off the Remainder. */
do while length(Remainder) > 1 & Left (Remainder,1) == '0'
Remainder = substr (Remainder, 2)
end
if Operator \== '/' then do
/* Check whether % would fail, even if operation is // */

if Floating() 'E' then do
if Operator '%' then MsgNum
else MsgNum

/* Page 133. % could fail by needing exponential notation */

26.11
26.12

call #Raise 'SYNTAX', MsgNum, Numberl , Number2, #Digits.#Level

end
end
if Operator == '//' then do
/* We need the remainder */
Sign3 = Signl

Mantissa3 = Remainder
Exponent3 = Exponentl - a
end

return /* From DivType */

/* The Power operation: */

Power: /* page 132 */
/* The second argument should be an integer */
if \WholeNumber2() then call #Raise 'SYNTAX', 26.8, Number2
/* Lhs to power zero is always 1 */
if Mantissa2 == '0' then do
Sign3 = '+'
Mantissa3
Exponent3
return
end

i
ror

/* Pages 132-133 The Power algorithm */
Rhs = left (Mantissa2,length(Mantissa2)+Exponent2,'0')/* Explicit
integer form */
L length (Rhs)
b X2B(D2X(Rhs)) /* Makes Rhs in binary notation */
/* Ignore initial zeros */
do q=l1byl
if substr(b,q,1) \== '0' then leave
end q
ael
do forever
/* Page 133 "Using a precision of DIGITS+L+1" */
if substr(b,q,1) == '1' then do
a = Recursion('*',Signl || Mantissal'E'Exponent1)

*/
if left(a,2) == 'MN' then signal PowerFailed
end

/* Check for finished */

if q = length(b) then leave

/* Square a */

a = Recursion('*',a)
if left(a,2) == 'MN' then signal PowerFailed
q=qrtil
end
/* Divide into one for negative power */
if Sign2 == '-' then do
Sign2 = '+'
a = Recursion('/')
if left(a,2) == 'MN' then signal PowerFailed
end

/* Split the value up so that PostOp can put it together with rounding */
Parse value Prepare(a,#Digits.#Level+L+1) with Sign3 Mantissa3 Exponent3
return

PowerFailed:
/* Distinquish overflow and underflow */
ReWas = substr(a,4)
if Sign2 = '-' then if ReWas == '42.1' then RcWas
else RcWas
call #Raise 'SYNTAX', RcWas, Numberl, '**', Number2
/* No return */

"42.2!
"42.1!

WholeNumber2:
numeric digits Digits
if #Form.#Level == 'S' then numeric form scientific

else numeric form engineering
return datatype (Number2, 'W')

Recursion: /* Called only from '**! */
numeric digits #Digits.#Level + L + 1
signal on syntax name Overflowed
/* Uses ArithOp again under new numeric settings. */
if arg(1) == '/' then return 1l/a
else return a * arg(2)
Over flowed:
return 'MN '.MN
\end{verbatim}

\hypertarget{functions}{%
\section{Functions}\label{functions}}

\hypertarget{invocation}{%
\subsection{Invocation}\label{invocation}}

@@

\hypertarget{evaluation-of-arguments}{%
\subsection{Evaluation of arguments}\label{evaluation-of-arguments}}

@@

\hypertarget{the-value-of-a-label}{%
\subsection{The value of a label}\label{the-value-of-a-label}}

@@

\hypertarget{the-value-of-a-function}{%
\subsection{The value of a function}\label{the-value-of-a-function}}

@@

\hypertarget{the-value-of-a-method}{%
\subsection{The value of a method}\label{the-value-of-a-method}}

@@

\hypertarget{the-value-of-a-message-term}{%
\subsection{The value of a message
term}\label{the-value-of-a-message-term}}

@@

/* This code is used in a context where \#Self is the receiver of the
method invocation which the subject message term is running under. */

\begin{verbatim}
SelectMethod:

/* If symbol given, receiver must be self. */
if arg(3,'E') then if arg(1)\==#Self then signal error /* syntax number? */

t arg(2) /* Will have been uppercased, unless a literal. */

x arg(1) /* Cursor through places to look for the method. */

Mixing 1 /* Off for potential mixins ignored because symbol given. */
Mixins -array~new /* to note any Mixins involved. */

/* Look in the method table of the object, if no 'symbol' given. */
if arg(3,'E') then do
Mixing = 0

end
else do
m = x~#MethodTable[t]
if m \== .nil then return m
end

do until x==.object
/* Follow the class hierarchy. */
x = x-class
/* Note any mixins for later reference. */
Mix = x~Inherited /* An array, ordered as the directive left-to-right. */

if Mix \== .nil then /* Append to the record. */
do j=1 to Mix~dimension (1)
Mixins [Mixins~dimension(1)+1] = Mix[j]
end

if Mixing do
/* Consider mixins only for superclasses of 'symbol'. */
do j=1 to Mixins~dimension (1)
/* Look at the baseclass of each. */
/* That is closest superclass not a mixin. */
s = Mixins[j]~class
do while s~Mixin /* Assert stop at .object if not before. */
s=s~class
end
if s==x then do
m=Mixins [j]~#MethodTable[t]
if m \== .nil then return m
end
end j
end /* Mixing */
if arg(3,'E') then if arg(3)==x then do
Mixing=1
end
if Mixing do
/* Consider non-Mixins */
m= x-#InstanceMethodTable[t]

if m \== .nil then return m
end

x=x~superclass

end

/* Try for UNKNOWN instead */
if t == 'UNKNOWN' then return .nil
if \arg(3,'E') then return SelectMethod arg(1),'UNKNOWN'
else return SelectMethod arg(1),'UNKNOWN',arg(3)
\end{verbatim}

\hypertarget{use-of-config_externalroutine}{%
\subsection{Use of
Config\_ExternalRoutine}\label{use-of-config_externalroutine}}

>>>>

%preprocessed texin
\chapter{Evaluation}\label{evaluation}

@@

\section{Variables}\label{variables}

@@

\subsection{Var\_Empty}\label{var_empty}

@@

\subsection{Var Set}\label{var-set}

@@

\subsection{Var\_Value}\label{var_value}

@@

\subsection{Var\_Drop}\label{var_drop}

@@

\subsection{Var\_Expose}\label{var_expose}

@@

\subsection{Var\_Reset}\label{var_reset}

@@

\section{Symbols}\label{symbols}

@@

\section{Value of a variable}\label{value-of-a-variable}

@@

\subsection{Value of a reserved
symbol}\label{value-of-a-reserved-symbol}

@@

\section{Expressions and operators}\label{expressions-and-operators}

@@

\subsection{The value of a term}\label{the-value-of-a-term}

See nnn for the syntax of a term.

@@

\subsection{The value of a
prefix\_expression}\label{the-value-of-a-prefix_expression}

@@

\subsection{The value of a
power\_expression}\label{the-value-of-a-power_expression}

@@

\lstinputlisting[language=rexx,label=evaluation-datatype.rexx,caption=evaluation-datatype.rexx]{evaluation-datatype.rexx}
power\_expression is

@@

\subsection{The value of a
multiplication}\label{the-value-of-a-multiplication}

@@

\subsection{The value of an addition}\label{the-value-of-an-addition}

@@

\subsection{The value of a
concatenation}\label{the-value-of-a-concatenation}

@@

\subsection{The value of a comparison}\label{the-value-of-a-comparison}

@@

If the comparison has a comparison\_operator that is a normal\_compare
then the variable \#Test is set as follows:
\lstinputlisting[language=rexx,label=evaluation-comparison.rexx,caption=evaluation-comparison.rexx]{evaluation-comparison.rexx}

\subsection{The value of an
and\_expression}\label{the-value-of-an-and_expression}

@@

\subsection{The value of an
expression}\label{the-value-of-an-expression}

@@

\subsection{Arithmetic operations}\label{arithmetic-operations}

@@

The code of ArithOp itself is assumed to operate under a sufficiently
high setting of numeric digits to avoid exponential notation.
\lstinputlisting[language=rexx,label=evaluation\_arithmetic.rexx,caption=evaluation\_arithmetic.rexx]{evaluation\_arithmetic.rexx}

\section{Functions}\label{functions}

\subsection{Invocation}\label{invocation}

@@

\subsection{Evaluation of arguments}\label{evaluation-of-arguments}

@@

\subsection{The value of a label}\label{the-value-of-a-label}

@@

\subsection{The value of a function}\label{the-value-of-a-function}

@@

\subsection{The value of a method}\label{the-value-of-a-method}

@@

\subsection{The value of a message
term}\label{the-value-of-a-message-term}

@@

/* This code is used in a context where \#Self is the receiver of the
method invocation which the subject message term is running under. */
\lstinputlisting[language=rexx,label=evaluation-selectmethod,caption=evaluation-selectmethod]{evaluation-selectmethod}

\subsection{Use of
Config\_ExternalRoutine}\label{use-of-config_externalroutine}

### 2024-02-16 ARB/standards/work-in-progress/document/tex/standard/instructions.tex
work in standard book production process including readme.md
upd

\hypertarget{instructions}{%
\chapter{Instructions}\label{instructions}}

@@

\hypertarget{routine-initialization}{%
\section{Routine initialization}\label{routine-initialization}}

@@

\hypertarget{clause-initialization}{%
\section{Clause initialization}\label{clause-initialization}}

@@

\hypertarget{clause-termination}{%
\section{Clause termination}\label{clause-termination}}

@@

\hypertarget{instruction}{%
\section{Instruction}\label{instruction}}

\hypertarget{address}{%
\subsection{ADDRESS}\label{address}}

@@

\hypertarget{assignment}{%
\subsection{Assignment}\label{assignment}}

@@

\hypertarget{call}{%
\subsection{CALL}\label{call}}

@@

\hypertarget{command-to-the-configuration}{%
\subsection{Command to the
configuration}\label{command-to-the-configuration}}

@@

\hypertarget{do}{%
\subsection{DO}\label{do}}

@@

\hypertarget{do-loop-tracing}{%
\subsection{DO loop tracing}\label{do-loop-tracing}}

@@

\hypertarget{drop}{%
\subsection{DROP}\label{drop}}

@@

\hypertarget{exit}{%
\subsection{EXIT}\label{exit}}

@@

\hypertarget{expose}{%
\subsection{EXPOSE}\label{expose}}

@@

\hypertarget{forward}{%
\subsection{FORWARD}\label{forward}}

@@

\hypertarget{guard}{%
\subsection{GUARD}\label{guard}}

@@

\hypertarget{interpret}{%
\subsection{INTERPRET}\label{interpret}}

@@

\hypertarget{iterate}{%
\subsection{ITERATE}\label{iterate}}

@@

\hypertarget{execution-of-labels}{%
\subsection{Execution of labels}\label{execution-of-labels}}

@@

\hypertarget{leave}{%
\subsection{LEAVE}\label{leave}}

@@

\hypertarget{message-term}{%
\subsection{Message term}\label{message-term}}

@@

\hypertarget{loop}{%
\subsection{LOOP}\label{loop}}

@@

\hypertarget{nop}{%
\subsection{NOP}\label{nop}}

@@

\hypertarget{numeric}{%
\subsection{NUMERIC}\label{numeric}}

@@

\hypertarget{numeric-digits}{%
\subsubsection{NUMERIC DIGITS}\label{numeric-digits}}

@@

\hypertarget{numeric-form}{%
\subsubsection{NUMERIC FORM}\label{numeric-form}}

@@

\hypertarget{numeric-fuzz}{%
\subsubsection{NUMERIC FUZZ}\label{numeric-fuzz}}

@@

\hypertarget{options}{%
\subsubsection{OPTIONS}\label{options}}

@@

\hypertarget{parse}{%
\subsubsection{PARSE}\label{parse}}

@@

\hypertarget{procedure}{%
\subsection{PROCEDURE}\label{procedure}}

@@

\hypertarget{pull}{%
\subsection{PULL}\label{pull}}

@@

\hypertarget{push}{%
\subsection{PUSH}\label{push}}

@@

\hypertarget{queue}{%
\subsection{QUEUE}\label{queue}}

@@

\hypertarget{raise}{%
\subsection{RAISE}\label{raise}}

@@

\hypertarget{reply}{%
\subsection{REPLY}\label{reply}}

@@

\hypertarget{select}{%
\subsection{SELECT}\label{select}}

@@

\hypertarget{signal}{%
\subsection{SIGNAL}\label{signal}}

@@

\hypertarget{trace-output}{%
\subsection{Trace output}\label{trace-output}}

@@

\hypertarget{conditions-and-messages}{%
\section{Conditions and Messages}\label{conditions-and-messages}}

@@

\hypertarget{raising-of-conditions}{%
\subsection{Raising of conditions}\label{raising-of-conditions}}

@@

\hypertarget{messages-during-execution}{%
\subsection{Messages during execution}\label{messages-during-execution}}

>>>>

%preprocessed texin
\chapter{Instructions}\label{instructions}

@@

\section{Routine initialization}\label{routine-initialization}

@@

\section{Clause initialization}\label{clause-initialization}

@@

\section{Clause termination}\label{clause-termination}

@@

\section{Instruction}\label{instruction}

\subsection{ADDRESS}\label{address}

@@

\subsection{Assignment}\label{assignment}

@@

\subsection{CALL}\label{call}

@@

\subsection{Command to the
configuration}\label{command-to-the-configuration}

@@

\subsection{DO}\label{do}

@@

\subsection{DO loop tracing}\label{do-loop-tracing}

@@

\subsection{DROP}\label{drop}

@@

\subsection{EXIT}\label{exit}

@@

\subsection{EXPOSE}\label{expose}

@@

\subsection{FORWARD}\label{forward}

@@

\subsection{GUARD}\label{guard}

@@

\subsection{INTERPRET}\label{interpret}

@@

\subsection{ITERATE}\label{iterate}

@@

\subsection{Execution of labels}\label{execution-of-labels}

@@

\subsection{LEAVE}\label{leave}

@@

\subsection{Message term}\label{message-term}

@@

\subsection{LOOP}\label{loop}

@@

\subsection{NOP}\label{nop}

@@

\subsection{NUMERIC}\label{numeric}

@@

\subsubsection{NUMERIC DIGITS}\label{numeric-digits}

@@

\subsubsection{NUMERIC FORM}\label{numeric-form}

@@

\subsubsection{NUMERIC FUZZ}\label{numeric-fuzz}

@@

\subsubsection{OPTIONS}\label{options}

@@

\subsubsection{PARSE}\label{parse}

@@

\subsection{PROCEDURE}\label{procedure}

@@

\subsection{PULL}\label{pull}

@@

\subsection{PUSH}\label{push}

@@

\subsection{QUEUE}\label{queue}

@@

\subsection{RAISE}\label{raise}

@@

\subsection{REPLY}\label{reply}

@@

\subsection{SELECT}\label{select}

@@

\subsection{SIGNAL}\label{signal}

@@

\subsection{Trace output}\label{trace-output}

@@

\section{Conditions and Messages}\label{conditions-and-messages}

@@

\subsection{Raising of conditions}\label{raising-of-conditions}

@@

\subsection{Messages during execution}\label{messages-during-execution}
